\textbf{\sc Towards Deliberative Control in Marine Robotics}

### 2012-04-30 europa.tex
added example to the architecture section.

\begin{figure}
\centering
\includegraphics[scale=0.4]{figs/EUROPA-Architecture.pdf}
\caption{\small The \eu software architecture.}
\label{fig:europa-architecture}
\end{figure}

@@

@@

This problem instance consists of \textbf{Variables} and
\textbf{Objects}, \textbf{Facts}, \textbf{Goals} and
\textbf{Constraints}. In our example we define temporal variables that
are used to specify the planning horizon (\texttt{horizonStart} and
\texttt{horizonEnd}) and objects are used to specify locations,
products, product-location associations and the Shopping Agent that we
want to plan for. Facts are used to specify the current state of the
world, do not need any support and are represented as tokens in the
\texttt{PDB}. In this case we use facts to specify that the Shipping
Agent is initially at home. Goals are also represented as tokens in
the \texttt{PDB}. Different from facts, goals trigger inference in \eu
which determines whether flaws (unresolved subgoals, overloaded
resources, etc.) need to be resolved; if so, search modules can be
invoked to resolve those flaws as part of the deliberation
process. Finally, constraints are restrictions on and between facts 
and goals. In the example above, we use temporal constraints to
specify when facts hold and when the goals should be
accomplished. In the model above, \texttt{before} and \texttt{after} 
are \texttt{Allen Algebra} \cite{allen84} operators, which \eu supports to specify
temporal relations between tokens (specifically between their start and end timepoints)
as summarized in Fig. \ref{fig:allen-algebra}

\begin{figure}[!htb]
\centering
\includegraphics[scale=0.3]{figs/Allen-algebra.pdf}
\caption{\small Temporal relations defined within the planner are
  based on \texttt{Allen Algebra} \cite{allen84} relations shown
  above.}
\label{fig:allen-algebra}
\end{figure}

@@

Since \eu supports a flexible temporal representation, where time is
represented by intervals, instead of single values, determining
whether a resource is the source of a plan flaw or a constraint
violation in general is substantially more difficult than just
computing a profile of the resource and checking against some limits.
With temporal flexibility, a number of resource profiles could be
possible depending on precisely when each production/consumption event
occurs out of a range of possible values that are in the domain of the
variables that represent the temporal extent for the event.
\cite{Muscettola04,Muscettola06} proposed an algorithm where the
variables that represent the time for production/consumption events
are placed in a temporal network, similar to the network used by the
temporal propagator; constraints on the resource can then be seen as
quantities flowing through that network. Computing the maximum flow
then, yields the maximum resource usage and thus a lower bound on the
resource level; computing a minimum flow will similarly yield an upper
bound. Since maximum flow on a temporal network can be computed in
polynomial time, \eu takes advantage of this algorithm to efficiently
compute upper and lower bound envelopes on the resource
level. Fig. \ref{fig:resenvelopes} shows an example of such a resource
computation.

\begin{figure}
\centering
\includegraphics[scale=0.2]{figs/europa-resource-envelopes.png}
\caption{\small An example of resource envelopes computed by \eu.}
\label{fig:resenvelopes}
\end{figure}

Resource envelopes can be useful when trying to evaluate plan
brittleness in the face of changing activity durations or start
times. They can also be useful when searching for plans; however they
are relatively expensive to compute. Therefore their ability to prune
the search space must be weighted against the ability to search faster
\cite{6007772}, \comment{cite "Stable Grounded Inference in Flexible Resource Scheduling" by Morris, Bresina, Barreiro}.

@@

Since temporal intervals may extend indefinitely in both positive and
negative directions, for problems with a temporal dimension it is also
common to specify a \emph{planning horizon} when invoking a solver;
any decisions that fall completely outside of the horizon can be
ignored. This is useful to ignore activities that may occur far in the future 
or in the past to be relevant for a particular plan, ensuring that the solver is making
as few commitments as necessary and reducing the complexity of plan generation.
The \emph{planning horizon} is also useful when dealing with recurrent activities;
consider for instance a model where a medical checkup has to occur
every $3$ months. The user can model a recurring activity and then use
the horizon to ensure that a finite number of medical checkup
activities are generated as part of a plan. 
\subsubsection{Flaw Resolution}
\label{sec:europa:flaws}

@@

\begin{algorithm}[H]
  \KwIn{ConstraintEngine $ce$}
  \KwIn{int $maxIter$}
  \Begin{
    int $curIter \leftarrow 0$ \;
    \While{$\left( ce.getViolations() > 0 \right) ~ \& ~ \left( curIter <
        maxIter \right)$ }{
      Variable $queenToMove \leftarrow getQueenWithMaxViolation(ce)$
      \;
      SortedSet$<$int$> moves \leftarrow getMoves(ce, queenToMove,
      curPos)$ \;
      bool $moved \leftarrow $ \texttt{false} \;
      \ForEach{int $newPos$ of $moves$}{
        $moved \leftarrow tryMove(queenToMove, curPos, newPos)$\;
        \lIf{$moved$}{ \textbf{break} \; }
      }
      \lIf{$! moved$}{
        $forceMove(queenToMove, curPos, moves.first())$ \;
      }
      $checkSolution()$ \tcp*{See if we have a new best solution}
      $curIter$\texttt{++} \;
    }
  }
  \caption{$solveNQueens(ce, maxIter)$}
\end{algorithm}

>>>>

\begin{figure}[b]
\centering
\includegraphics[scale=0.5]{figs/EUROPA-Architecture.pdf}
\caption{\small The \eu software architecture.\comment{Is there a
    better block diagram we can use to respond to review comment on
    page 2?}}
\label{fig:europa-architecture}
\end{figure}

@@

Before diving into the details of each module, a simple example illustrates how they interact at a high level:

class LightBulb extends Timeline
{
    predicate On  {}
    predicate Off {}
}

class LightSwitch extends Timeline
{
    LightBulb myBulb_;

    LightSwitch(LightBulb b)
    {
        myBulb_ = b;
    }

    action turnOn  { duration=1; }
    action turnOff { duration=1; }
}

LightSwitch::turnOn
{
    met_by(condition object.myBulb_.Off);  // Bulb must be Off to be turned On
    meets(effect object.myBulb_.On);   // Must be turned on through the switch
}

LightSwitch::turnOff
{
    met_by(condition object.myBulb_.On);   // Bulb must be On to be turned Off
    meets(effect object.myBulb_.Off);      // Must be turned off through the switch
}

// Problem instance : turning the light off
LightBulb bulb1 = new LightBulb();
LightSwitch switch1 = new LightSwitch(bulb1);

// At time 0, the bulb is on
fact(bulb1.On initialCondition);
eq(initialCondition.start,0);

// We want the bulb to be off by time 10
goal(bulb1.Off goal1);
lt(0,goal1.start);
lt(goal1.start,10);

The snippet above is a NDDL model of a light bulb and a switch that controls it (NDDL is covered in more detail below), along with a problem instance that requires the light bulb to be off by time=10, given an initial state with the bulb on at time=0. The NDDL module allows \eu to ingest this model and problem description. Once the NDDL statements are parsed, the different object types, object instances and constraints are instantiated inside the Plan Database and Constraint Engine modules where they are available for both the user and the solver module to query and manipulate programmatically. This instantiation of the initial state constitutes the initial plan that \eu holds for this particular problem instance. 
The user can then invoke functionality available in the solver module to modify or extend the initial plan in a way that eliminates any flaws (violated constraints, unsatisfied goals, oversubscribed resources) in the current plan. When actions are added to a plan, the Rules Engine module ensures that domain rules expressed in the model are enforced when \eu creates new action or predicate instances, for instance, if a LightSwitch::turnOn action is added to the plan, the RulesEngine  ensures that requirements for the existence of the conditions and effects of that action (that is, the corresponding LightBulb must be off before and on after the action) are posted through the PlanDatabase and ConstraintEngine modules.

@@

This problem instance consists of \textbf{Variables} and
\textbf{Objects}, \textbf{Facts}, \textbf{Goals} and
\textbf{Constraints}. In our example we define temporal variables that
are used to specify the planning horizon (\texttt{horizonStart} and
\texttt{horizonEnd}) and objects are used to specify locations,
products, product-location associations and the Shopping Agent that we
want to plan for. Facts are used to specify the current state of the
world, do not need any support and are represented as tokens in the
\texttt{PDB}. In this case we use facts to specify that the Shipping
Agent is initially at home. Goals are also represented as tokens in
the \texttt{PDB}. Different from facts, goals trigger inference in \eu
which determines whether flaws (unresolved subgoals, overloaded
resources, etc.) need to be resolved; if so, search modules can be
invoked to resolve those flaws as part of the deliberation
process. Finally, constraints are restrictions on and between facts
and goals. In the example above, we use temporal constraints to
specify when facts hold and when the goals should be accomplished. In
the model above, \texttt{before} and \texttt{after} are \texttt{Allen
  Algebra} \cite{allen84} operators, which \eu supports to specify
temporal relations between tokens (specifically between their start
and end timepoints). The algebra is particularly effective in dealing
with interval arithmetic; its elements are summarized in
Fig. \ref{fig:allen-algebra}

\begin{figure}[!t]
\centering
\includegraphics[scale=0.4]{figs/Allen-algebra.pdf}
\caption{\small Temporal relations defined within the planner are
  based on \texttt{Allen Algebra} \cite{allen84} relations shown
  above.}
\label{fig:allen-algebra}
\end{figure}

@@

Since \eu supports a flexible temporal representation, where time is
represented by intervals, instead of single values, determining
whether a resource is the source of a plan flaw or a constraint
violation in general is substantially more difficult than just
computing a profile of the resource and checking against some limits.
With temporal flexibility, a number of resource profiles could be
possible depending on precisely when each production/consumption event
occurs out of a range of possible values that are in the domain of the
variables that represent the temporal extent for the event.
\cite{Muscettola04,Muscettola06} proposed an algorithm where the
variables that represent the time for production/consumption events
are placed in a temporal network, similar to the network used by the
temporal propagator; constraints on the resource can then be seen as
quantities flowing through that network. Computing the maximum flow
then, yields the maximum resource usage and thus a lower bound on the
resource level; computing a minimum flow will similarly yield an upper
bound. Since maximum flow on a temporal network can be computed in
polynomial time \comment{definition needed?}, \eu takes advantage of
this algorithm to efficiently compute upper and lower bound envelopes
on the resource level. Fig. \ref{fig:resenvelopes} shows an example of
such a resource computation.

\begin{figure}
\centering
\includegraphics[scale=0.3]{figs/europa-resource-envelopes.png}
\caption{\small An example of resource envelopes computed by \eu.}
\label{fig:resenvelopes}
\end{figure}

Resource envelopes can be useful when trying to evaluate plan
brittleness in the face of changing activity durations or start
times. They can also be useful when searching for plans; however they
are relatively expensive to compute. Therefore their ability to prune
the search space must be weighted against the ability to search faster
\cite{6007772, morris11}.

@@

Since temporal intervals may extend indefinitely in both positive and
negative directions, for problems with a temporal dimension it is also
common to specify a \emph{planning horizon} when invoking a solver;
any decisions that fall completely outside of the horizon can be
ignored. This is useful to ignore activities that may occur far in the
future or in the past to be relevant for a particular plan, ensuring
that the solver is making as few commitments as necessary and reducing
the complexity of plan generation.  The \emph{planning horizon} is
also useful when dealing with recurrent activities; consider for
instance a model where a medical checkup has to occur every $3$
months. The user can model a recurring activity and then use the
horizon to ensure that a finite number of medical checkup activities
are generated as part of a plan.

\subsubsection{Flaw Resolution}
\label{sec:europa:flaws}

@@

\begin{algorithm}[H]
  \KwIn{ConstraintEngine $ce$}
  \KwIn{int $maxIter$}
  \Begin{
    int $curIter \leftarrow 0$ \;
    \While{$\left( ce.getViolations() > 0 \right) ~ \& ~ \left( curIter <
        maxIter \right)$ }{
      Variable $queenToMove \leftarrow getQueenWithMaxViolation(ce)$
      \;
      SortedSet$<$int$> moves \leftarrow getMoves(ce, queenToMove,
      curPos)$ \;
      bool $moved \leftarrow $ \texttt{false} \;
      \ForEach{int $newPos$ of $moves$}{
        $moved \leftarrow tryMove(queenToMove, curPos, newPos)$\;
        \lIf{$moved$}{ \textbf{break} \; }
      }
      \lIf{$! moved$}{
        $forceMove(queenToMove, curPos, moves.first())$ \;
      }
      $checkSolution()$ \tcp*{Evaluate if a new best solution}
      $curIter$\texttt{++} \;
    }
  }
  \caption{$solveNQueens(ce, maxIter)$}
\end{algorithm}
